### 2018-07-17 submission/manuscript.Rmd
update discussion and methods

Previous studies have reported GC dependent amplification differences [@polz_bias_1998] and we confirm that these differences still persist in high fidelity DNA polymerase when analyzing mock community samples [Figure 1]. Additionally, these specific polymerases have different sequence error rates and chimera prevalence [Figure 2]. Chimera prevalence in particular was strongly associated with the total number of OTUs detected in a mock community [Figure 2E]. This total number of OTUs also varied by polymerase used for these mock community samples [Figure 3A]. The range in the number of OTUs observed increased with the number of cycles and the change in range was dependent on the polymerase for both mock community and fecal samples [Figure 3]. However, these differences did not translate to large bacterial community composition changes based on polymerase [Figure 4]. Even though there were bacterial community composition changes based on the number of cycles they occurred independently of the polymerase used [Figure 4]. Additionally, there were no OTU-based differences observed within the number of cycles between any of the polymerases for fecal samples. When Random Forest models were created to classify number of cycles, polymerase, and individual based on the 16S rRNA gene sequencing, the model performance for individuals was significantly greater than all other models tested [Figure 5A]. In addition to the better model performance of individuals, the most important OTUs to this model were not the most important OTUs for the other models tested [Figure 5B]. Our observations show that differences between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences are dwarfed by biological differences between fecal bacterial communities within individuals.

@@

***Analysis workflow.***
To adjust for unequal sequencing, all samples were rarefied to 1000 sequences for downstream analysis. Analysis of the mock community of each polymerase for GC-based amplification differences, sequence error rate, number of sequences with an error, base substitution, and numbers of chimeras before and after chimera removal with VSEARCH was assessed. Additionally, the correlation between the number of chimeras and the number of OTUs was also assessed. The total number of OTUs, taxa differences, and Bray-Curtis indices were analyzed for both the fecal and mock community samples. Finally, Random Forest models were created to assess whether classification of polymerase, cycles, or individual could be performed best using 16S rRNA gene sequencing data. Additionally, overlap between the most important OTUs to the three models was assessed using mean decrease in accuracy (MDA).

>>>>

Previous studies have reported GC dependent amplification differences [@polz_bias_1998] and we confirm that these differences still persist in high fidelity DNA polymerase when analyzing mock community samples [Figure 1]. Additionally, these specific polymerases have different sequence error rates and chimera prevalence [Figure 2]. Chimera prevalence in particular was strongly associated with the total number of OTUs detected in our mock community and vaired by polymerase  [Figure 2E & 3A]. Additionally, the bacterial community composition varied by number of cycles and polymerase in these mock community samples. Clearly, it is possible to identify a large number of differences based on number of cycles and polymerase used when analyzing mock community samples. However, these differences may not be a major concern when analyzing more complex community samples.

Fecal samples showed a similar increase in median and range for the number of OTUs, based on number of cycles and polymerase used, as that found when analyzing mock community samples [Figure 4A]. However, there were few differences between the polymerases within each of the amplification cycles [Figure 4A]. Additionally, these differences did not translate to large bacterial community composition changes based on number of cycles or polymerase. The differences that were identifed were small and occured based on the number of cycles, independent of the polymerase used [Figure 4B]. There was also no OTU-based differences observed within the number of cycles between any of the polymerases for fecal samples. Interestingly, Random Forest models built on this OTU data had the highest classification success for individuals versus number of cycles or polymerase. In addition to the better model performance of Random Forest for individuals, the most important OTUs to this model were not the most important OTUs for the other models tested [Figure 5]. Collectively, these observations show that differences based on the number of cycles and between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences can be smaller than the biological differences between fecal bacterial communities within individuals.

@@

***Analysis workflow.***
To adjust for unequal sequencing, all samples were rarefied to 1000 sequences for downstream analysis. The analysis of the mock community of each polymerase for GC-based amplification differences, sequence error rate, number of sequences with an error, base substitution, and numbers of chimeras before and after chimera removal with VSEARCH was assessed. Additionally, the correlation between the number of chimeras and the number of OTUs was also assessed. The total number of OTUs, taxa differences, and Bray-Curtis indices were analyzed for both the fecal and mock community samples. Finally, Random Forest models were created to assess whether classification of polymerase, number of cycles, or individual could be performed best using 16S rRNA gene sequencing data. For the model used to classify individuals, all samples were included from all number of cycles and polymerases but only the individual labels were used. Both number of cycles and polymerase models included all samples but only the number of cycles or polymerase label was used for each respective model. Additionally, overlap between the most important OTUs to the three models was assessed using mean decrease in accuracy (MDA).

### 2018-07-17 submission/manuscript.tex
update discussion and methods

Previous studies have reported GC dependent amplification differences
(Polz \& Cavanaugh, 1998) and we confirm that these differences still
persist in high fidelity DNA polymerase when analyzing mock community
samples {[}Figure 1{]}. Additionally, these specific polymerases have
different sequence error rates and chimera prevalence {[}Figure 2{]}.
Chimera prevalence in particular was strongly associated with the total
number of OTUs detected in a mock community {[}Figure 2E{]}. This total
number of OTUs also varied by polymerase used for these mock community
samples {[}Figure 3A{]}. The range in the number of OTUs observed
increased with the number of cycles and the change in range was
dependent on the polymerase for both mock community and fecal samples
{[}Figure 3{]}. However, these differences did not translate to large
bacterial community composition changes based on polymerase {[}Figure
4{]}. Even though there were bacterial community composition changes
based on the number of cycles they occurred independently of the
polymerase used {[}Figure 4{]}. Additionally, there were no OTU-based
differences observed within the number of cycles between any of the
polymerases for fecal samples. When Random Forest models were created to
classify number of cycles, polymerase, and individual based on the 16S
rRNA gene sequencing, the model performance for individuals was
significantly greater than all other models tested {[}Figure 5A{]}. In
addition to the better model performance of individuals, the most
important OTUs to this model were not the most important OTUs for the
other models tested {[}Figure 5B{]}. Our observations show that
differences between high fidelity DNA polymerases can be found. However,
when using 16S rRNA gene sequencing these technical differences are
dwarfed by biological differences between fecal bacterial communities
within individuals.

@@

\textbf{\emph{Analysis workflow.}} To adjust for unequal sequencing, all
samples were rarefied to 1000 sequences for downstream analysis.
Analysis of the mock community of each polymerase for GC-based
amplification differences, sequence error rate, number of sequences with
an error, base substitution, and numbers of chimeras before and after
chimera removal with VSEARCH was assessed. Additionally, the correlation
between the number of chimeras and the number of OTUs was also assessed.
The total number of OTUs, taxa differences, and Bray-Curtis indices were
analyzed for both the fecal and mock community samples. Finally, Random
Forest models were created to assess whether classification of
polymerase, cycles, or individual could be performed best using 16S rRNA
gene sequencing data. Additionally, overlap between the most important
OTUs to the three models was assessed using mean decrease in accuracy
(MDA).

>>>>

Previous studies have reported GC dependent amplification differences
(Polz \& Cavanaugh, 1998) and we confirm that these differences still
persist in high fidelity DNA polymerase when analyzing mock community
samples {[}Figure 1{]}. Additionally, these specific polymerases have
different sequence error rates and chimera prevalence {[}Figure 2{]}.
Chimera prevalence in particular was strongly associated with the total
number of OTUs detected in our mock community and vaired by polymerase
{[}Figure 2E \& 3A{]}. Additionally, the bacterial community composition
varied by number of cycles and polymerase in these mock community
samples. Clearly, it is possible to identify a large number of
differences based on number of cycles and polymerase used when analyzing
mock community samples. However, these differences may not be a major
concern when analyzing more complex community samples.

Fecal samples showed a similar increase in median and range for the
number of OTUs, based on number of cycles and polymerase used, as that
found when analyzing mock community samples {[}Figure 4A{]}. However,
there were few differences between the polymerases within each of the
amplification cycles {[}Figure 4A{]}. Additionally, these differences
did not translate to large bacterial community composition changes based
on number of cycles or polymerase. The differences that were identifed
were small and occured based on the number of cycles, independent of the
polymerase used {[}Figure 4B{]}. There was also no OTU-based differences
observed within the number of cycles between any of the polymerases for
fecal samples. Interestingly, Random Forest models built on this OTU
data had the highest classification success for individuals versus
number of cycles or polymerase. In addition to the better model
performance of Random Forest for individuals, the most important OTUs to
this model were not the most important OTUs for the other models tested
{[}Figure 5{]}. Collectively, these observations show that differences
based on the number of cycles and between high fidelity DNA polymerases
can be found. However, when using 16S rRNA gene sequencing these
technical differences can be smaller than the biological differences
between fecal bacterial communities within individuals.

@@

\textbf{\emph{Analysis workflow.}} To adjust for unequal sequencing, all
samples were rarefied to 1000 sequences for downstream analysis. The
analysis of the mock community of each polymerase for GC-based
amplification differences, sequence error rate, number of sequences with
an error, base substitution, and numbers of chimeras before and after
chimera removal with VSEARCH was assessed. Additionally, the correlation
between the number of chimeras and the number of OTUs was also assessed.
The total number of OTUs, taxa differences, and Bray-Curtis indices were
analyzed for both the fecal and mock community samples. Finally, Random
Forest models were created to assess whether classification of
polymerase, number of cycles, or individual could be performed best
using 16S rRNA gene sequencing data. For the model used to classify
individuals, all samples were included from all number of cycles and
polymerases but only the individual labels were used. Both number of
cycles and polymerase models included all samples but only the number of
cycles or polymerase label was used for each respective model.
Additionally, overlap between the most important OTUs to the three
models was assessed using mean decrease in accuracy (MDA).
